---
title: "Dung Beetle Microbiome Figures"
format: html
editor: visual
---

## Dung beetle microbiome response to drenches

**Hypotheses**

1.  The use of drenches changes the dung beetles' microbiome composition

2.  Dung beetles' microbiome changes with diet and it is more "volatile than the microbiome of other insects

3.  Species richness of gut microbiome is lower than species richness from the rest of the dung beetle (gut vs whole)

## Method notes

**Treatments**

-   Before: Dung beetles directly from the rearing facility -- from North Island, had different dung in their diet than in our study

-   Control: Dung beetles that were not subjected to any drench in their diet during the experiment

-   Drenches low: A known concentration of drenches was applied to the non-drenched dung and then fed to the beetles

-   Drenches high: 10x concentration of the drenches low treatment was applied to the non-drenched dung and then fed to the beetles

Drenches used:

-   Silver: 5 5 g/l Moxidectin

-   Blue: 2g/l abamectin, 80g/l levamisole and 45.3g/l oxfendazole

-   Red: 80g/l levamisole and 45.3g/l oxfendazole

    Key messages from the paper:

-   Diet is the main factor in their microbiome composition

-   Due to their changing microbiome nature, they are perfect candidate for microbiome engineering studies

```         
Known drench degraders:
```

-   Stenotrophomonas -- Family Xanthomonadaceae -- Class Gammaproteobacteria

-   Aeromonas -- Family Aeromonadaceae -- Class Gammaproteobacteria

```         
Presumed drench degraders:
```

-   Pseudomonas -- Family Pseudomonadaceae -- Class Gammaproteobacteria

    **Additional question: Gammaproteobacteria are known drench degraders - are they more prevalent in the gut of the dung beetle when exposed to drenches?**

```{r}
# packages
library(phyloseq)
library(tidyverse)
library(readxl)
library(decontam)

# load data
tax_df <- read_xlsx("Dung beetle microbiome paper/db_classified_16S.xlsx") %>% 
  select(ASV,domain, phylum, class, order,family, genus)
otu_df <- read_xlsx("Dung beetle microbiome paper/otu_frequency_table.xlsx")
metadata_df <- read_xlsx("Dung beetle microbiome paper/Dung_beetles_metadata.xlsx") %>% 
  separate(Treatment, c("Treatment_base", "level"), remove = F) %>% 
  mutate(level = if_else(is.na(level), "NA", level)) %>% 
  rename(Extraction = Exctraction)
```

```{r}
# make phyloseq object
otumat <- as.matrix(otu_df[-1])
rownames(otumat) <- otu_df$ASV

taxmat<- as.matrix(tax_df[-1])
rownames(taxmat) <- tax_df$ASV



OTU <- otu_table(otumat, taxa_are_rows = TRUE)
TAX <- tax_table(taxmat)
SAM <- sample_data(metadata_df)
rownames(SAM) <- metadata_df$Samplename

ps <- phyloseq(OTU, TAX)
ps <- merge_phyloseq(ps, SAM)

```

```{r}
# check library sizes

df <- as.data.frame(sample_data(ps)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Treatment)) + geom_point()+
  facet_grid(level~Treatment_base)


```

Library size looks good! The negative and wash samples have very few reads. And the others look like they have a similar spread.

```{r}
# clean the data using the control sample "Wash" and "Negative controls"

sample_data(ps)$is.neg <- sample_data(ps)$Decontamination == "Control_sample"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold = 0.5) ## increase the threshold from 0.1 to 0.5
table(contamdf.prev$contaminant) ## 18 contaminants found

ps.noncontam_bac <- prune_taxa(!contamdf.prev$contaminant, ps) ## drop these 18 contaminants

DATA = ps.noncontam_bac  %>% 
  subset_samples(Decontamination != "Control_sample") # drop these controls and just keep the real samples that are now "clean"

```

```{r}


# selected 15,000 based on the library depths above; drops 13 samples and 60 OTUs
## could also use other methods for normalisation!
DATA_rarefied = rarefy_even_depth(physeq=DATA,sample.size=15000,rngseed=7,replace=FALSE,trimOTUs=TRUE,verbose=TRUE) 

DATA_rarefied <- prune_taxa(taxa_sums(DATA_rarefied)>0, DATA_rarefied)
  

```

**Hypotheses**

1.  The use of drenches changes the dung beetles' microbiome composition

```{r}
# just the drench and the control samples - exclude the "before" samples
DATA_experiment <- DATA_rarefied %>% 
  subset_samples(Treatment != "Before") # drop these

DATA_experiment <- prune_taxa(taxa_sums(DATA_experiment)>0, DATA_experiment)
  

ord <- ordinate(DATA_experiment, "MDS", "bray")

plot_ordination(DATA_experiment, ord, color = "Treatment_base", shape = "level") +
  geom_point(size = 3)+
  scale_color_manual(values = c( "blue2",
                                  "black", "red2",
                                 "grey70"))+
  theme_test()+
  facet_wrap(~Extraction)
    

psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}

# 

df_OTU <- psotu2veg(DATA_experiment) ## taxa are now columns and rows are samples, this is the correct format for vegan

df_metadata <- data.frame(sample_data(DATA_experiment)) %>% 
  unite(trt_extraction, c("Treatment", "Extraction"), sep = "_", remove = F) %>% 
  unite(trt_base_extraction, c("Treatment_base", "Extraction"), sep = "_", remove = F)

library(vegan)
dist.mat_all <- vegdist(df_OTU, method = "bray", binary=F)



OTU_distmat_all <-
  as.matrix(dist.mat_all, labels = T)

adon.results_all<-vegan::adonis2(OTU_distmat_all ~ Treatment_base*Extraction*level, data = df_metadata, method="bray",perm=999) ## Treatment, extraction and their interaction are all significant

adon.results_all
## do pairwise analysis to see which groups are different

## pairwise permanova
library(pairwiseAdonis)
pairwise.adonis(OTU_distmat_all,df_metadata$trt_base_extraction, p.adjust.m ="bonferroni")# -- interesting pairwise comparisons here, control is never different from the treatments -- could suggest that the treatments cause the gut microbiome to become more variable, but that the control communities are still persisting -- trts "add" to the control community?


# 
# perm <- how(nperm = 199)
# setBlocks(perm) <- with(metadata_df, sample)
# blocked_adonis_away <- adonis2(OTU_distmat_all_away ~ time.point*treat, data = metadata_df_away, permutations = perm) ## 
# 
# blocked_adonis_away

```

2.  Dung beetles' microbiome changes with diet and it is more "volatile than the microbiome of other insects

3.  Species richness of gut microbiome is lower than species richness from the rest of the dung beetle (gut vs whole)
